\textbf{{\Large {Recursive algorithms, branching coefficients affine algebras and applications}}}

### 2010-01-12 AffineLieAlgebras/InformalNotes/article.tex
Proof for $\mathfrak{a}_{\bot}$ using case-by-case analysis.

In Weyl group terms to construct $\Delta _{\mathfrak{a}_{\bot}} $ means the following: given a Weyl group $W$ and a subgroup $W_{\mathfrak{a}}$ find the subgroup $W_{\mathfrak{a}_{\bot}}\subset W$ such that $\omega\, W_{\mathfrak{a}}\,\omega^{-1}=W_{\mathfrak{a}} \;\text{for}\; \omega\in W_{\mathfrak{a}_{\bot}}$, so we need to find the normalizer {\color{red} (not the centralizer?, no proof)} of $W_{\mathfrak{a}}$ in $W$, (such that some of the elements of $W_{\mathfrak{a}_{\bot}}$ are the reflections corresponding to the positive roots of $\mathfrak{g}$  {\color{red}(?) and if so where is the proof?}). This task can be carried out using the computational methods of finite group theory \cite{stembridge2001computational} (but we are unaware of any publication with the exhaustive tables of such subgroups {\color{red}(drop it in the final text?)}).

\begin{comment}
  We need to proof the following:
  For the Weyl group $W_{\mathfrak{a}}$ any element $\omega in W$ such that $\omega\cdot W_{\mathfrak{a}} \cdot \omega = W_{\mathfrak{a}}$ has the form $\omega=r_1^{s_1} r_2^{s_2}\dots r_k^{s_k}$ where $r_1\dots r_k$ are the generators corresponding to the nodes of non-connected component if $\mathfrak{g}=A_n, C_n, D_n$. 
  For $\mathfrak{g}=B_n$ there exist additional elements $\{\omega=r_1^{s_1} r_2^{s_2}\dots r_k^{s_k}\cdot r_{k+1}^{s_{k+1}} r_{k+2}^{s_{k+2}}\}$. 
\end{comment}

For the series $A_n$ the statement has the following form:
\begin{statement}
  Consider subalgebra $\mathfrak{a}$ with the simple roots $\{\alpha_i, i=1\dots k,\; k<n\}$. Then every root $\alpha\bot \beta\in \Delta_{\mathfrak{a}}$ is the linear combination of the simple roots $\alpha_{k+2}\dots \alpha_n$ or equivalently in the language of Weyl groups, for $\forall \omega\in W: \omega\cdot W_{\mathfrak{a}}\cdot \omega=W_{\mathfrak{a}},\; \omega=r_{k+2}^{s_{k+2}}\dots r_n^{s_n}$, where $\{r_i\}$ are the Weyl reflections corresponding to simple roots $\{\alpha_i\}$.

\end{statement}
  \begin{proof}
    Assume the opposite. It is clear that it is enough to check that $r_i\cdot \omega\cdot r_i\neq \omega,\; i=1\dots k$.  
  \end{proof}

>>>>

For the algebra $\mathfrak{g}$ from the series $A_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r+1\}$ are $\Delta=\{\alpha_{ i j}= e_i- e_j,\; 1\leq i, j\leq r+1\}$, $\Delta^{+}=\{\alpha_{ij},\; i<j\}$ and the set of simple roots consists of $\alpha_{1,2},\alpha_{2,3},\dots,\alpha_{r,r+1}$. So for the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consisting of first $r_{\mathfrak{a}}$ simple roots we get $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{ i j},\; r_{\mathfrak{a}}+1<i,j\leq r+1\}$ and $\mathfrak{a}_{\bot}=A_{r-r_{\mathfrak{a}}-1}$.

For the algebra $\mathfrak{g}$ from the series $B_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r\}$ are $\Delta=\{\alpha_{\pm i,\pm j}=\pm e_i\pm e_j,\;i<j;\; \alpha_{\pm j}=\pm e_{ j},\; 1\leq j\leq r\}$, $\Delta^{+}=\{\alpha_{i,-j},\; \alpha_{ij},\; \alpha_j;\; i<j,\; 1\leq j\leq r\}$ and the set of simple roots consists of $\alpha_{1,-2},\alpha_{2,-3},\dots,\alpha_{r-1,-r},\alpha_r$. So if the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consists of first $r_{\mathfrak{a}}$ simple roots, then $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; r_{\mathfrak{a}}+1<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}-1}$. Otherwise if $\mathfrak{a}=B_{r_{\mathfrak{a}}}$ and its simple roots are $\alpha_{r-r_{\mathfrak{a}}+1,-r+r_{\mathfrak{a}}-2},\dots,\alpha_{r-1,r},\alpha_r$ we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=B_{r-r_{\mathfrak{a}}}$. It is the only case when simple roots of $\mathfrak{a}_{\bot}$ can not be obtained from the extended Dynkin diagram, as can be seen in the Table \ref{tab:diagrams}.

For the algebra $\mathfrak{g}$ from the series $C_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r\}$ are $\Delta=\{\alpha_{\pm i,\pm j}=\pm e_i\pm e_j,\;i<j;\; \alpha_{\pm j}=\pm 2e_{ j},\; 1\leq j\leq r\}$, $\Delta^{+}=\{\alpha_{i,-j},\; \alpha_{ij},\; \alpha_j;\; i<j,\; 1\leq j\leq r\}$ and the set of simple roots consists of $\alpha_{1,-2},\alpha_{2,-3},\dots,\alpha_{r-1,-r},\alpha_r$. So if the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consists of first $r_{\mathfrak{a}}$ simple roots, then $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; r_{\mathfrak{a}}+1<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=C_{r-r_{\mathfrak{a}}-1}$. Otherwise if $\mathfrak{a}=C_{r_{\mathfrak{a}}}$ and its simple roots are $\alpha_{r-r_{\mathfrak{a}}+1,-r+r_{\mathfrak{a}}-2},\dots,\alpha_{r-1,r},\alpha_r$ we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\;\alpha_j,\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=C_{r-r_{\mathfrak{a}}}$.

For the algebra $\mathfrak{g}$ from the series $D_r$ the roots in the orthogonal basis $\{e_i,\; 1\leq i\leq r\}$ are $\Delta=\{\alpha_{\pm i,\pm j}=\pm e_i\pm e_j,\;1\leq i<j\leq r\}$, $\Delta^{+}=\{\alpha_{i,-j},\; \alpha_{ij},\; i<j,\; 1\leq j\leq r\}$ and the set of simple roots consists of $\alpha_{1,-2},\alpha_{2,-3},\dots,\alpha_{r-1,-r},\alpha_{r-1,r}$. So if the regular subalgebra $\mathfrak{a}=A_{r_{\mathfrak{a}}}$ and its simple root system consists of first $r_{\mathfrak{a}}$ simple roots, then $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\; r_{\mathfrak{a}}+1<i<j\leq r\}$ and $\mathfrak{a}_{\bot}=D_{r-r_{\mathfrak{a}}-1}$. Otherwise if $\mathfrak{a}=D_{r_{\mathfrak{a}}}$ and its simple roots are $\alpha_{r-r_{\mathfrak{a}}+1,-r+r_{\mathfrak{a}}-2},\dots,\alpha_{r-1,r},\alpha_{r-1,r}$ we see that $\Delta_{\mathfrak{a}_{\bot}}=\{\alpha_{\pm i,\pm j},\; 1<i<j\leq r-r_{\mathfrak{a}}\}$ and $\mathfrak{a}_{\bot}=D_{r-r_{\mathfrak{a}}}$.
